---
title: "Q & K Box 14.1"
output: 
  html_notebook:
    theme: flatly
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Deriving components (modified from Jackson 2003)
There are two strategies for extracting eigenvectors (components) and their eigenvalues from a multivariate data set of n objects by p variables. First, we can use an eigenanalysis (spectral decomposition) of a p x p association matrix between variables. Second, we can use a singular value decomposition (SVD) of the n x p data matrix, with variables standardized as necessary.

Consider the matrix (Y) of raw data from Clevenger and Waltho (2000), who recorded the numbers of seven taxa of wildlife for eleven underpasses in Alberta, Canada.

Load the vegan package
```{r results='hide'}
library(tidyverse)
library(vegan)
```

```{r}
#datasets: clevenger.csv
clevenger <- read_csv("../data/clevenger.csv")
head(clevenger,10)
```
### Illustrate centroid using relationship between human activity index (hui) and elk numbers
Plots used for QK use the ggplot classic theme, with some tweaks. Tweaks are consolidated into theme_QK; use this theme for figures and tweak the theme to avoid repetitive code changes.
```{r echo=FALSE, results='hide'}
source("../R/appearance.R")
```
Fig 14.2
```{r}
f14_2<-ggplot(data=clevenger,aes(x=hui, y=elk))+
  geom_point(color=sc, size=ss, alpha= 1) +
  theme_qk()+
  geom_point(aes(x=mean(hui),y=mean(elk)), pch=8, size=3)
f14_2
```

### Get centered and standardized raw data matrices
```{r}
# centered
clevenger.cen <- scale(clevenger[,c('bbear','gbear','cougar','wolf','deer','elk','moose')], scale=F)
clevenger.cen
# get standardized raw data
clevenger.stan <- scale(clevenger[,c('bbear','gbear','cougar','wolf','deer','elk','moose')], scale=T)
clevenger.stan
```
### Generate covariance and correlation matrices
Covariance matrix for wildlife
```{r}
clevenger.cov <- cov(clevenger[,c('bbear','gbear','cougar','wolf','deer','elk','moose')])
```
Correlation matrix for wildlife
```{r}
clevenger.cor <- cor(clevenger[,c('bbear','gbear','cougar','wolf','deer','elk','moose')])
```

### Derive eigenvectors and eigenvalues based on correlation matrix
```{r}
clevenger.eigen <- eigen(clevenger.cor)
clevenger.eigen
```
### Do PCA
```{r}
clevenger.pca <- princomp(~bbear+gbear+cougar+wolf+deer+elk+moose, cor=T, data=clevenger)
summary(clevenger.pca)
```
### Calculate dissimilarities for raw and standardized (by object and variables) and presence/absence
Calculate Euclidean, Chisq and Bray-Curtis measures
```{r}
# raw data
clev1 <- clevenger[c(16:22)]
vegdist(clev1, method="euclidean")
vegdist(clev1, method="chisq")
vegdist(clev1, method="bray")
# standardized by object totals
clev1$total <- rowSums(clev1)
clev2 <- clev1
clev2$bbeart <- clev2$bbear/clev2$total
clev2$gbeart <- clev2$gbear/clev2$total
clev2$cougart <- clev2$cougar/clev2$total
clev2$wolft <- clev2$wolf/clev2$total
clev2$deert <- clev2$deer/clev2$total
clev2$elkt <- clev2$elk/clev2$total
clev2$mooset <- clev2$moose/clev2$total
clev3 <- clev2[c(9:15)]
vegdist(clev3, method="euclidean")
vegdist(clev3, method="chisq")
vegdist(clev3, method="bray")
# standardized by variable totals
clev1$bbeart <- clev1$bbear/sum(clev2$bbear)
clev1$gbeart <- clev1$gbear/sum(clev2$gbear)
clev1$cougart <- clev1$cougar/sum(clev2$cougar)
clev1$wolft <- clev1$wolf/sum(clev2$wolf)
clev1$deert <- clev1$deer/sum(clev2$deer)
clev1$elkt <- clev1$elk/sum(clev2$elk)
clev1$mooset <- clev1$moose/sum(clev2$moose)
clev4 <- clev1[c(9:15)]
vegdist(clev4, method="euclidean")
vegdist(clev4, method="chisq")
vegdist(clev4, method="bray")
```
**create pres/abs datafile**
```{r error=TRUE}
clev5 <- clevenger[c(16:22)]
clev5[clev5>0] <-1
clev5
vegdist(clev5, method="jaccard")
vegdist(clev5, method="bray")
```
Check outliers with Mahalanobis distances - no particularly unusual objects
```{r}
# Problem with code; needs to be fixed
#mahalanobis(clev1, colMeans(clev1), cov(clev1))
```

### Fig 14.3 - ordination with BC MDS
```{r}
clev6 <- clevenger[c(1,16:22)]
clev6.stand <- decostand(clev6[,2:8], "total")
clev6.stand.bc <- vegdist(clev6.stand,"bray")
clev6.mds <- metaMDS(clev6.stand.bc,k=2)
plot(clev6.mds$points,type="n", )
text(clev6.mds$points,lab=clev6$underpass, col="black")
```
```{r, error=TRUE}
mdsplot <- as.data.frame(scores(clev6.mds))
mdsplot$underpass<-clev6$underpass
f14_3<-ggplot(data=mdsplot,aes(x=NMDS1, y=NMDS2, label=underpass))+
  geom_text()+
  theme_qk()+
  theme(axis.text = element_blank())
f14_3
```
