---
title: "QK Box 2.2"
output:
  html_document:
    df_print: paged
  html_notebook:
    theme: flatly
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}
# This code block sets up the r session when the page is rendered to html

# Change general options
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, 
                      message = FALSE, warning = FALSE, 
                      cache = FALSE, par = FALSE)

# Set up figure options
knitr::opts_chunk$set(fig.align = 'center', fig.show = 'asis', 
                      fig.width = 3.5, fig.height = 4) 

# Set the web address where R will look for files from this repository
repo_url <- "https://raw.githubusercontent.com/elahi/qk2_examples/master/"

# Set working directory
knitr::opts_knit$set(root.dir = "../")
```

Low et al (2016) examined the effects of two different anesthetics on aspects of the physiology of the mouse. Twelve mice were anesthetized with isoflurane and eleven mice were anesthetized with alpha chloralose and blood CO~2~ levels were recorded after 120 minutes. The *H*~0~ was that there is no difference between the anesthetics in the mean blood CO~2~ level. This is an independent comparison because individual mice were only given one of the two anesthetics.

## Preliminaries

First, load the required packages (tidyverse, RMisc, MKinfer, car, emmeans)

```{r include=FALSE, results='hide'}
# source("../R/libraries.R")
# source("../R/appearance.R")

source("R/libraries.R")
source("R/appearance.R")
```

Import low data file

```{r}
low <- read.csv("data/lowco2.csv")
low
```

## Get summary statistics by anesthetic

```{r }
low_stats <- summarySE(data=low,measurevar="co2", groupvars="anesth")
low_stats
low %>% dplyr::count(anesth)
low %>%  
  group_by(anesth) %>% 
  dplyr::summarise(median = median(co2),
            mean = mean(co2))
```

## Plot data

```{r}
low %>% 
  ggplot(aes(anesth, co2)) + 
  geom_point(alpha = 0.5) + 
  theme_qk()
```

## Fit model and get effect size

```{r}
low.aov <- aov(co2~anesth,data=low)
tidy(low.aov, conf.int=TRUE)

low.emm <- emmeans(low.aov,"anesth")
eff_size(low.emm, sigma=sigma(low.aov), edf=df.residual(low.aov))
```

Note that we've chosen to show a standardized effect size, using the pooled variance from the analysis of variance - Residual MS = 262.44, and √262.44 = 16.2

## Test variances

```{r}
leveneTest(co2 ~ anesth, low)
```

### t-test for equal variances

```{r}
t.test(co2~anesth,var.equal=TRUE, data=low)
```

### t-test for separate variances

```{r}
t.test(co2~anesth,data=low)
```

## Wilcoxon-Mann-Whitney

```{r}
wilcox.test(co2~anesth,data=low)
sum(rank(low$co2)[low$anesth=="ac"])
sum(rank(low$co2)[low$anesth=="iso"])
```
