---
title: "Q & K Box 13.3"
output:
    html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Calculations in ecotoxicology. The data used here are summary vaules taken from the Bliss data used by Finney in his description of probit analysis. Data are in 2 forms - individual live/dead values, and summary proportions for each dose.

The data files are [here](../data/findisagg.csv) and [here](../data/finneyagg.csv)

### Preliminaries

First, load the required packages (vcd)

```{r }
source("../R/libraries.R")   #This is the common library
library(performance)
library(vcd)
```

## Disaggregated data file

```{r}
finndisagg <- read_csv("../data/finneydisagg.csv")
head(finndisagg)
plot(dead~dose, data=finndisagg)
```

Fit GLM with binomial response variable

```{r}
fdisagg.lm<-glm(dead~dose,family=binomial, data=finndisagg)
tidy(fdisagg.lm, conf.int=TRUE)
plot(fdisagg.lm$fitted.values~finndisagg$dose)
coef(fdisagg.lm)
summary(fdisagg.lm) #get model fit
```

Calc LC50 Use parameter values from the dose-response, solve for 50%

```{r }
3.28/.612
```

### Original analysis used log 10 transform of dose

```{r}
fdisagg.lm3<-glm(dead~ldose,family=binomial, data=finndisagg, na.action=na.omit)
tidy(fdisagg.lm3, conf.int=TRUE)
summary(fdisagg.lm3) #get model fit
```

solve for y=0; use coefficients to calculate LC50

```{r }
coef(fdisagg.lm3)
5.12/7.43
```

backtransform to get LC50 on raw dose scale

```{r }
10^0.689
```

### Fit same model as probit, rather than logit

```{r }
fdisagg.lm4<-glm(dead~ldose,family=binomial(link="probit"), data=finndisagg, na.action=na.omit)
tidy(fdisagg.lm4, conf.int=TRUE)
summary(fdisagg.lm4) #get model fit
```

Use coefficients to calculate LC50

```{r }
coef(fdisagg.lm4)
3.03/4.38
10^0.692
```

### For comparison to original analyses, we'll use aggregated values, generate probit and logit values for each dose and fit linear regression.

```{r }
finney <- read_csv("../data/finneyagg.csv")
head(finney)
```

#### Fit logit model

```{r}
finn.lm<-lm(logit~l10dose, data=finney, na.action=na.omit)
tidy(finn.lm)
#Use coefficients to calculate LC50

coef(finn.lm)
4.805/7.004


# Back-transform to untransformed dose

10^.686
```

#### Probit transformed values

```{r }
finn.lm2<-lm(probit_raw~l10dose,data=finney, na.action=na.omit)
tidy(finn.lm2)
# use coefficients to calculate LC50
coef(finn.lm2)
2.85/4.15
# Back-transform
10^0.687
```

**Note**: Finney used probits with 5 added

```{r}
finn.lm3<-lm(probit_finney~l10dose,data = finney, na.action=na.omit)
tidy(finn.lm3)
# solve for y=5
(5-2.14)/4.16
10^.688
```
