---
title: "QK Box 16.1"
output:
  html_notebook
---

title: "Q & K Box 16.1" output: html_notebook ---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Worked example of PCoA: invertebrates in artificial ponds This example continues with the pond invertebrate example used in Boxes [15.2](lemmensCA.nb.html) and [15.3](lemmensRDA.nb.html) and uses the data file [lemminvert.csv](../data/lemminvert.csv).

### Preliminaries

Load core set of packages and vegan, ape

```{r results='hide'}
source("../R/libraries.R")
library(vegan)
library(ape)
```

Read data

```{r}
lemminvert <- read.csv("../data/lemminvert.csv")
```

## Do PCoA on lemmens invert data

Use B-C matrix on raw abundances

```{r}
# Remove pond classification variables
lemminvert.bc <- vegdist(lemminvert[,4:32],method="bray")
# do PCoA
lemminvert.pco <- pcoa(lemminvert.bc)
print(lemminvert.pco)
biplot(lemminvert.pco,lab=lemminvert$manag)
```

### Improve plot

Plots used for QK use the ggplot classic theme, with some tweaks. Tweaks are consolidated into theme_QK; use this theme for figures and tweak the theme to avoid repetitive code changes.

```{r}
source("../R/appearance.R")
```

Get labels back on

```{r}
# extract coordinates
a<-as.data.frame(lemminvert.pco$vectors)
# Need to get management types on this plot
a<-cbind(lemminvert[c(1:3)],a)   #Add site names & symbols from original data file
br=c("nm","li","nf","yf")     #Categories from original file; define as object to avoid retyping
la=c("None", "Light", "No fish", "Young fish")   #Labels; define here to avoid retyping
p2<-ggplot(data=a, aes(x=Axis.1, y=Axis.2, shape=manag, ) )+
  geom_point()+
  labs(y="MDS2", x="MDS1")+
  scale_shape_manual(values=sym4,
                     name="Management",
                     breaks=br,
                     labels=la,
                     guide =
                         guide_legend(label.theme = element_text(size=6),
                                    title=NULL)
                     )+
  theme_qk()+
  theme(legend.position = "top")
p2
```

```{r}
# colour version using viridis palette d
p2a<-ggplot(data=a, aes(x=Axis.1, y=Axis.2, color=manag, ) )+
  geom_point()+
  labs(y="MDS2", x="MDS1")+
  scale_color_viridis_d(
                     name="Management",
                     breaks=br,
                     labels=la,
                     guide =
                         guide_legend(label.theme = element_text(size=6),
                                    title=NULL)
                     )+
  theme_qk()+
  theme(legend.position = "top")
p2a
```
