---
title: "QK Box 9.1"

output: 
  html_notebook:
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This box continues with the forest bird data set used in [Boxes 8.2](loyn.nb.html) and [8.10](loyncat.nb.html)

We calculated common measures of relative importance based on the fit of the following full model relating bird abundance to five predictors for the dataset from Loyn (1997; see Box 8.2): (bird abundance)~i~ = β~0~ + *β*~1~(log patch area) + *β*~2~(log distance to nearest patch) +*β*~3~(grazing) +*β*~4~(altitude) + *β*~5~(year isolated) + *ε~i~*

### Preliminaries

First, load the required packages (relaimpo, car, hier.part, MuMIn, lm.beta, sjstats)

```{r include=FALSE, results='hide', error=TRUE}
source("R/libraries.R")
library(hier.part)      #This package has been removed from CRAN
library(relaimpo)
library(lm.beta)
library(sjstats)
```

Import loyn data file ([data/loyn.csv](data/loyn.csv)); from previous example, we needed to transform area and distance

```{r}
loyn <- read.csv("../data/loyn.csv")
head(loyn,10)
loyn$logarea <- log10(loyn$area)
loyn$logdist <- log10(loyn$dist)
```

For some analyses, we also need two dataframes that are subsets of loyn containing the response variable and the predictors

```{r}
loyn_abund<-loyn$abund
loyn_pred<-subset(loyn, select = c("logarea","logdist","alt","yearisol","graze"))
```

### Run model from Box 8.2

```{r }
loyn.lm <- lm(abund~ logarea+logdist+graze+alt+yearisol, data=loyn)
tidy(loyn.lm)
```

#### Get Standardized coefficients (usual)

```{r }
lm.beta.loyn <- lm.beta(loyn.lm)
tidy(lm.beta.loyn, conf.int=TRUE)
```

#### Standardized coefficients (both usual and partial sd)

```{r error=FALSE}
std.coef(loyn.lm, partial.sd=FALSE)
std.coef(loyn.lm, partial.sd=TRUE)
```

## Get relative importance

```{r}
calc.relimp(loyn.lm, type = c("lmg", "pmvd", "last", "first", "betasq", "pratt"), rela=FALSE)
```

**Note:** Available metrics from *relaimpo* package: lmg, pmvd (non-US version only), last, first, betasq, pratt, genizi and car. - from package help

Can only get pmvd to run if we obtain the non-US version from author's website and install manually - copy into relevant R library folder.

Author web site is <http://prof.beuth-hochschule.de/groemping/software/relaimpo/>

**Note:** Use stand coefficents based on partial sd above instead of betasq

```{r }
loyn.boot <- boot.relimp(loyn.lm, b=1000, type = c("lmg", "pmvd"))
booteval.relimp(loyn.boot)
```

### Now look at hierarchical partitioning

```{r error=TRUE}
hier.part(loyn_abund, loyn_pred, family="gaussian", gof="Rsqu")
```

The package hier.part was removed from CRAN in March 2023. The code above will work if you have hier.part installed already.
An alternative is to use the package *glmm.hp*, which is done in the next code chunk.

Hier.part can also be installed from Github, though there may be issues with M1/M2 Macs. The quick way from Github is using devtools:
devtools::install_github("cjbwalsh/hier.part")

```{r}
library (glmm.hp)
glmm.hp(loyn.lm, type="R2")
```

### Model selection (unstandardized coefficients)

```{r }
options(na.action = "na.fail")
loyn.dredge <-dredge(loyn.lm, beta="none", evaluate=TRUE)
loyn.dredge
```

### Model selection (standardized coefficients, by partial sd)

```{r }
options(na.action = "na.fail")
loyn.dredge1 <-dredge(loyn.lm, beta="partial.sd", evaluate=TRUE)
loyn.dredge1
```

## Model averaging

**Note** confint only provides CIs for conditional (natural) averages, based on z=1.96. Modified SE CIs in Box 9.1 calculated by hand for full averaging

### Unstandardized first

```{r }
loyn.ma<-model.avg(loyn.dredge)
summary(loyn.ma)
confint(loyn.ma)
```

### Get standardized (by partial sd and unconditional sd) model averaged estimates

```{r error=TRUE}
loyn.ma1<-model.avg(loyn.dredge, beta="sd") #Recent update gives error with this averaging
summary(loyn.ma1)
confint(loyn.ma1)
loyn.ma2<-model.avg(loyn.dredge, beta="partial.sd")
summary(loyn.ma2)
confint(loyn.ma2)
```

### Sum of akaike weights

```{r }
sw(loyn.dredge)
# importance(loyn.dredge)   #this function now defunct, and sw recommended
```

## k-fold cross validation

```{r }
cv_error(loyn, abund~ logarea+logdist+graze+alt+yearisol, k=5)
cv_error(loyn, abund~ logarea+logdist+graze+alt+yearisol, k=10)
```

### leave one out cross-validation

```{r }
cv_error(loyn, abund~ logarea+logdist+graze+alt+yearisol, k=56)
```
