---
title: "QK Box 8.11"

output: 
  html_notebook:
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We re-analysed the data from Loyn (1987; see Box 8.2 ) by fitting a simpler model that just included grazing and log patch area (based on our model selection criteria -- see Box 9.1). First we treated grazing as a continuous variable like we did in the original analysis by fitting the following model:

(bird abundance)~*i*~ = *β~0~* + *β~1~*(log~10~ area)~*i*~ + *β*~2~(grazing)~*i*~ *+ ε~i~*

### Preliminaries

First, load the required packages (car, lm.beta)

```{r include=FALSE, results='hide', error=TRUE}
source("../R/libraries.R")   #This is the common library
library(lm.beta)
```

Import [loyncat](data/loyncat.csv) data file

This file is the loyn data, with grazing classified into five categories ("grazecat"). We could have simply turned the numerical grazing level into a factor, but this way the grazing levels have meaningful names

```{r}
loyncat <- read.csv("../data/loyncat.csv")
head(loyncat,10)
loyncat$grazecat<-factor(loyncat$grazecat)
```

### Fit linear model with grazing as continuous variable

```{r }
loyncat1.lm <- lm(abund~log10(area)+graze, data=loyncat)
tidy(loyncat1.lm, conf.int=TRUE)
```

Get standardized coefficients

```{r }
lm.beta(loyncat1.lm)
```

### Fit linear model with grazing as a categorical predictor

Use zero grazing as the reference category

```{r }
loyncat2.lm <- lm(abund~log10(area)+relevel(grazecat,ref="zero"), data=loyncat)
tidy(loyncat2.lm, conf.int=TRUE)
```

Check residuals - look fine

```{r }
plot(loyncat2.lm)
```

Get standardised coefficients

```{r }
lm.beta(loyncat2.lm)
```

### Get added variable plots

```{r }
avPlots(loyncat2.lm, ask=F)
```
