---
title: "QK Box 10.6"
output:
  html_document:
    df_print: paged
  html_notebook:
    theme: flatly
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, 
                      warning = FALSE, message = FALSE)

# Set working directory
knitr::opts_knit$set(root.dir = "../")

library(knitr)
```

Caballes et al. (2016) examined the effects of maternal nutrition (three treatments: starved or fed one of two coral genera: *Acropora* or *Porites*) on the larval biology of crown-of-thorns seastars. There were three female seastars nested within each treatment, 50 larvae reared from each female were placed into each of three glass culture jars and the lengths of ten larvae from each jar after four days were measured after 4 days. This fully balanced design has maternal nutrition as a fixed factor with three random levels of nesting: females within nutrition treatment, jars within females and individual larvae within jars.

![Crown of thorns seastar, Acanthaster planci. Mick Keough, [CC BY 4.0](https://creativecommons.org/licenses/by-sa/4.0)](../media/Crown_of_thorns.jpg){width="686"}

![Larval cycle of crown-of-thorns. [AIMS](www.aims.gov.au), [CC BY 4.0](https://creativecommons.org/licenses/by-sa/4.0)](../media/Larval%20cycle%20of%20crown-of-thorns.png)

The paper is [here](doi:10.1371/journal.pone.0158007)

Caballes, C. F., Pratchett, M. S., Kerr, A. M. & Rivera-Posada, J. A. (2016). The role of maternal nutrition on oocyte size and quality, with respect to early larval development in the coral-eating starfish, *Acanthaster planci*. *PLoS One*, 11, e0158007.

### Preliminaries

Packages:

```{r}
#| results: hide
library(tidyverse)
source("R/set_ggplot_theme.R")
library(broom)
library(Rmisc)
library(car)
library(lme4)
library(lmerTest)
library(nlme)
library(VCA)
library(afex)
```

Data:

```{r}
caballes_length <- read.csv("data/caballes_length.csv")
```

```{r }
set_sum_contrasts()
```

Make female and jar factors

```{r }
caballes_length$female <- factor(caballes_length$female)
caballes_length$jar <- factor(caballes_length$jar)
```

Reorder nutrition treatments so starved is first for default lm contrasts

```{r }
caballes_length$diet <- factor(caballes_length$diet, levels = c("starved","aa","pr"))
levels(caballes_length$diet)
```

## Visualize data

```{r}
caballes_length %>% 
  ggplot(aes(jar, length, color = female)) + 
  geom_point(alpha = 0.5) +
  facet_wrap(~ diet, scales = "free_x")
```

## Run model

**Note:** can't get the aov commands to work for 3 level nested design

```{r error=TRUE, eval = FALSE}
caballes.aov <- aov(length~diet+Error(female/jar), caballes_length)
```

### Fit as lm using OLS estimation

```{r }
caballes.lm <- lm(length ~ diet/female/jar, caballes_length)
anova(caballes.lm)
```

Check diagnostics and lm output (not shown):

```{r}
#| eval: false
plot(caballes.lm)
summary(caballes.lm)
```

Get F and P values using correct denominators

```{r }
#Diet F
f <- 1.26668/0.06224
pf(f, df1 = 2, df2 = 6, lower.tail = FALSE)
#Females F
f <- 0.06224/0.02925
pf(f, df1 = 6, df2 = 18, lower.tail = FALSE)
#Jars F
f <- 0.02925/0.01755
pf(f, df1 = 18, df2 = 243, lower.tail = FALSE)
```

Variance components from VCA

```{r}
#| eval: false
caballes.vca <- anovaMM(length ~ diet/(female)/(jar), caballes_length)
caballes.vca
VCAinference(caballes.vca, alpha = 0.05, VarVC = TRUE, ci.method = "satterthwaite")
```

### Fit mixed effects model using REML/ML

```{r }
caballes.lmer <- lmer(length ~ diet + (1|female/jar), caballes_length)
summary(caballes.lmer)
```

Get F-ratio for diet test using lmerTest

```{r }
anova(caballes.lmer, ddf = "Kenward-Roger")
```

CI on variance components (remembering to square CIs from lmer which are in SD units)

```{r }
caballes.ci <- confint.merMod(caballes.lmer)
caballes.vc <- (caballes.ci)^2
print(caballes.vc)
```

## Simplify dataset by averaging across larvae within a jar

```{r}
d <- caballes_length %>% 
  group_by(diet, female, jar) %>%
  dplyr::summarise(mean = mean(length), 
            sd = sd(length), 
            n = n(), 
            se = sd / sqrt(n)) %>% 
  ungroup()

d %>% 
  ggplot(aes(female, mean)) + 
  geom_point(alpha = 0.5) +
  facet_wrap(~ diet, scales = "free_x") + 
  labs(y = "Mean length (mm)")
```

### Fit nested ANOVA with OLS

Female is nested within treatment:

```{r}
d_aov <- aov(mean~diet+Error(female), d)
summary(d_aov)

m1 <- lm(mean ~ diet / female, d)
anova(m1)
```

Get F and P values using correct denominators. Note that I'm using `broom:tidy` to tidy the `anova` output, and `lead()` to calculate the `new_F` and `new_P`. This avoids extracting `statistic` (F-value) and `df` into new objects, or hard-coding the calculations.

```{r }
tidy(anova(m1)) %>% 
  mutate(new_F = meansq / lead(meansq), 
         new_P = pf(new_F, df1 = df, df2 = lead(df), lower.tail = FALSE)) %>% 
  kable(digits = 3)
```

### Fit mixed effects model using REML/ML

```{r }
m2 <- lmer(mean ~ diet + (1|female), d)
summary(m2)
anova(m2, ddf = "Kenward-Roger")
```
