---
title: "QK Box 17.4"
output:
  html_notebook:
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We illustrate how “default” error bars produced by most software packages can be misleading when summarizing mixed models. We generated two simple data sets for a three-factor mixed model design. Factor A has two levels and is fixed. Factor B is a random effect with four levels and is nested in A. Factor C is crossed with A & C and is fixed with three levels. In the first data set (response variable Y1), there are strong effects of the random factor, and effects of factor C are consistent across that factor. The second data set (Y2) has identical observations for each combination of A & C, but data were permuted so the effects of C vary across levels of B. 

### Preliminaries

First, load the required packages (afex, car, lattice, lme4, lmerTest, nlme, VCA, ez, emmeans) and, for convenience, apaTables

```{r include=FALSE, results='hide'}
source("../R/libraries.R")
source("../R/appearance.R")
library(apaTables)
```

Import errbar data file (errbar.csv)

```{r}
errbar <- read.csv("../data/errbar.csv")
errbar
```

set contrasts from afex
make individual a factor, make sure species a factor too

```{r results='hide'}
set_sum_contrasts()
errbar$B <- factor(errbar$B)
```


### Fit full model with Y1

```{r }
errbar1.aov <- aov(Y1~A*C+Error(B),errbar)
summary(errbar1.aov)
```

use ez for comparison with type 3 SS - same result as design is balanced

```{r }
ezerrbar <- ezANOVA(data=errbar, dv=Y1, wid=B, within=C, between=A, type=3)
print(ezerrbar)
#GG issue, so print clean table with corrections
apa.ezANOVA.table(ezerrbar)
```

### Repeat for Y2
```{r }
errbar2.aov <- aov(Y2~A*C+Error(B),errbar)
summary(errbar2.aov)
```

### Graphics
Plot raw data for Y1 and Y2
```{r}
graphY1<-ggplot(data=errbar, aes(x=C, y=Y1, color=A, group=B))+
    geom_point(size=2,show.legend = FALSE)+
  geom_line(show.legend=FALSE)+
  theme_qk()+scale_color_viridis(discrete=TRUE, option="H", begin=0.25)
graphY2<-ggplot(data=errbar, aes(x=C, y=Y2, color=A, group=B))+
    geom_point(size=2,show.legend = FALSE)+
  geom_line(show.legend=FALSE)+
  theme_qk()+scale_color_viridis(discrete=TRUE, option="H", begin=0.25)
emm1<-emmeans(errbar1.aov, ~A|C)      #extract means
emm2<-as.data.frame(emm1)                    #convert data table to frame
```
Plot means and standard errors
The plot uses Y1, but could as easily be done with Y2
```{r}
pd=position_dodge(width=0.1)
graphmean<-ggplot(emm2,aes(x=C,y=emmean,shape=A, group=A, color=A))+
  geom_point(size=3,show.legend = FALSE, position=pd)+    #size=3 for larger symbols, as means shown (so fewer points on fig)
  geom_errorbar(aes(ymin = emmean-SE, ymax = emmean+SE), position=pd, width=0, show.legend = FALSE)+
  geom_line(show.legend=FALSE, position=pd)+
  labs(y="Mean")+
   theme_qk()+scale_color_viridis(discrete=TRUE, option="H", begin=0.25)
```
Combined graph
```{r}
graphY1+graphmean+graphY2
```


