---
title: "QK Box 11.2"
output:
  html_notebook:
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Legault et al. (2018) experimentally examined the effects of nutrient loading on the outcome of competitive interactions between two species of saltmarsh plants, the native *Spartina alterniflora* and the exotic *Phragmites australis*, in northeast USA. The between-plots factor was nutrient addition (fixed, with three groups: none, low and high) with five 50-gallon bins containing half-strength seawater as the plots. The within-plots factor was competition (fixed with two groups: no competition, with competition), with sub-plots being small pots containing one (no competition) or both (with competition) species within each bin. This design was replicated at the sub-plot level with two pots for each competition group within each bin. We will analyze biomass of *P. australis* at ambient temperature as the response variable.

![*Phragmites australis*. Mick Keough, [CC BY 4.0](https://creativecommons.org/licenses/by-sa/4.0)](../media/phragmites2.jpg){width="512"}

The paper is [here](https://doi.org/10.1371/journal.pone.0100718)

Legault, R., Zogg, G. P. & Travis, S. E. (2018). Competitive interactions between native *Spartina alterniflora* and non-native *Phragmites australis* depend on nutrient loading and temperature. *PLoS One*, 13, e0192234.

### Preliminaries

First, load the required packages (afex, car, lattice, lme4, lmerTest, nlme, VCA, ez, emmeans, Rmisc, MuMIn)

```{r include=FALSE, results='hide'}
source("../R/libraries.R")   #This is the common library
```

Import legault data file ([legault.csv](data/legault.csv))

```{r}
legault <- read.csv("../data/legault.csv")
legault
```

Set contrasts from afex

Re-arrange nutrient levels so that no nutrients is reference

```{r }
set_sum_contrasts()
legault$nutrient <- factor(legault$nutrient, levels=c("Null (0 g N/m2/year)","Low (30 g N/m2/year)","High (120 g N/m2/year)"))
legault$comp<-factor(legault$comp)
```

### Fit aov crossed model to check approx residuals

```{r }
legault1.aov <- aov(biom~nutrient+comp+nutrient*comp, legault)
plot(legault1.aov)
```

Some evidence for unequal residuals but use untransformed biomass to match original paper

### Fit model using aov (balanced design so all SS types the same)

```{r }
legault2.aov <- aov(biom~nutrient+Error(bin/comp)+comp+nutrient*comp, legault)
summary(legault2.aov)
print(summary(legault2.aov))
```

### Fit split-plot anova model using ez for comparison of output

This uses mean biomass for each competition group within each bin - same answers as above due to balanced design

```{r }
legault.ez <- ezANOVA(dv=biom, wid=bin, within=comp, between=nutrient, type=1, legault)
print(legault.ez)
```

### Rerun as classic split-plot design based on sub-plot means - matches above

This example illustrates a short-cut that produces a simpler analysis for a paper, etc.

```{r }
legault_means <- summarySE(legault, measurevar= 'biom', groupvars= c('nutrient','bin','comp'))
legault_means
legault3.aov <- aov(biom~nutrient*comp+Error(bin), legault_means)
summary(legault3.aov)
```

### Fit mixed model using lme4

```{r }
legault1.lmer <- lmer(biom~nutrient+comp+nutrient*comp+(1|bin)+(1|bin:comp), REML=TRUE, legault)
```

Get residuals from lmer model fit - some evidence on variance heterogeneity but again use untransformed biom to match original paper

```{r }
plot(legault1.lmer)
summary(legault1.lmer, ddf="Kenward-Roger")
anova(legault1.lmer, ddf="Kenward-Roger")
```

Contrast no nutrients with low and high separately

```{r }
legault1.emm <- emmeans(legault1.lmer, ~nutrient)
legault1.con <- contrast(legault1.emm, "trt.vs.ctrl", ref=1)
summary(legault1.con, adjust="none")
```

Get variance components with CIs

```{r }
legault1.ci <- confint.merMod(legault1.lmer, oldNames=FALSE)
legault1.vc <- (legault1.ci)^2
print(legault1.vc)
```

Evaluate whether to omit bins(nutrients) x competition interaction

```{r error=TRUE}
legault1a.lmer <- lmer(biom~nutrient+comp+nutrient*comp+(1|bin)+(1|bin:comp), REML=FALSE, legault)
legault2.lmer <- lmer(biom~nutrient+comp+nutrient*comp+(1|bin), REML=FALSE, legault)
anova(legault1a.lmer,legault2.lmer)
AICc(legault1a.lmer,legault2.lmer)
```

Get results from model omiting bins(nutrients) x competition interaction

```{r error=TRUE }
legault2a.lmer <- lmer(biom~nutrient+comp+nutrient*comp+(1|bin), REML=TRUE, legault)
summary(legault2a.lmer)
anova(legault2a.lmer, type="3", ddf="Kenward-Roger")
```

### Fit mixed model on means (classic split-plot design) using lmer

```{r }
legault3.lmer <- lmer(biom~nutrient+comp+nutrient*comp+(1|bin), REML=TRUE, legault_means)
summary(legault3.lmer)
anova(legault3.lmer, ddf="Kenward-Roger")
```

Get variance components with CIs

```{r }
legault3.ci <- confint.merMod(legault3.lmer)
legault3.vc <- (legault3.ci)^2
print(legault3.vc)
```
