---
title: "QK Box 16.2"
output:
  html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Worked example of NMDS, ANOSIM, PERMANOVA, MV-ABUND, BIO-ENV: invertebrates in artificial ponds

We continue with the pond invertebrate example from [Box 16.1](lemmpcoa.nb.html). We will now use these data to produce an NMDS ordination plot showing the relationship among ponds based on invertebrate family abundances. We will compare this analysis to one where the abundances are converted to proportions of pond totals and where abundances are converted to presence-absence data. We will also compare the four management groups using ANOSIM, PERMANOVA, and MV-ABUND and examine relationships between the patterns among ponds based on invertebrate families and patterns based on environmental variables.

### Preliminaries 

Load graphics tweaks and core set of packages

```{r results='hide', echo=FALSE}
#Plots used for QK use the ggplot classic theme, with some tweaks. Tweaks are consolidated into theme_QK; use this theme for figures and tweak the theme to avoid repetitive code changes.
source("../R/appearance.R")
source("../R/libraries.R")
```

Load additional packages: vegan, mvabund

```{r results='hide', echo=FALSE}
library(vegan)
library(mvabund)
```

Load lemminvert data and remove labels

```{r results='hide'}
lemminvert <- read_csv("../data/lemminvert.csv")
lemminvert1 <- lemminvert[,-(1:3)]
```

### MDS on original invert abundances

```{r}
lemminvert.bc <- vegdist(lemminvert1,'bray')    #Create Bray-Curtis matrix
lemminvert.mds <- metaMDS(lemminvert.bc,k=2,autotransform=FALSE)
stressplot(lemminvert.mds, main="Shepard plot")
lemminvert.mds
plot(lemminvert.mds$points,type="n")
text(lemminvert.mds$points,lab=lemminvert$manag,col="black")
points(lemminvert.mds$points,pch=16,col=as.numeric(lemminvert$managsymb))
```

### Standardize abundance to proportional abundance for each pond

```{r}
lemminvert1s <- decostand(lemminvert1, method = "total")
lemminvert1s.bc <- vegdist(lemminvert1s,'bray')
lemminvert1s.mds <- metaMDS(lemminvert1s.bc,k=2,try=20,trymax=40,maxit=200,autotransform=FALSE)
stressplot(lemminvert1s.mds, main="Shepard plot")
lemminvert1s.mds
plot(lemminvert1s.mds$points,type="n")
text(lemminvert1s.mds$points,lab=lemminvert$manag,col="black")
points(lemminvert1s.mds$points,pch=16,col=as.numeric(lemminvert$managsymb))
```

### MDS on presence-absence using Jaccard transformed data

```{r}
lemminvert.jac <- vegdist(lemminvert1,binary=TRUE,method='jaccard')
lemminvert.jac.mds <- metaMDS(lemminvert.jac,k=2,autotransform=FALSE)
stressplot(lemminvert.jac.mds, main="Shepard plot")
lemminvert.jac.mds
plot(lemminvert.jac.mds$points,type="n")
text(lemminvert.jac.mds$points,lab=lemminvert$manag,col="black")
points(lemminvert.jac.mds$points,pch=16,col=as.numeric(lemminvert$managsymb))
```

### Plot env variables on ordinations

Select 11 variables (same as ch15)

```{r error=TRUE}
lemmenv <- read_csv("../data/lemmenv.csv")
lemmenv1 <- lemmenv[,c(3:5,7:8,10:13,16:17)]
# standardize predictors
lemmenv1$lchla <- log10(lemmenv1$chla)
lemmenv1$depth <- scale(lemmenv1$depth)
lemmenv1$silt <- scale(lemmenv1$silt)
lemmenv1$ph <- scale(lemmenv1$ph)
lemmenv1$temp <- scale(lemmenv1$temp)
lemmenv1$surface <- scale(lemmenv1$surface)
lemmenv1$cond <- scale(lemmenv1$cond)
lemmenv1$lchla <- scale(lemmenv1$lchla)
lemmenv1$tn <- scale(lemmenv1$tn)
lemmenv1$tp <- scale(lemmenv1$tp)
lemmenv1$reed <- scale(lemmenv1$reed)
lemmenv1$emerg <- scale(lemmenv1$emerg)
# plot using original abundances
fit1 <- envfit(lemminvert.mds, lemmenv1)
fit1
plot(fit1)
# plot using proportional abundances
fit2 <- envfit(lemminvert1s.mds, lemmenv1)
fit2
plot(fit2)
# plot using pa
fit3 <- envfit(lemminvert.jac.mds, lemmenv1)
fit3
plot(fit3)
```

### Model/test management types

Check homogeneity of dispersions on original abundances; if OK, for comparison run anosim, permanova and mvabund

```{r}
lemminvert.disp <- betadisper(lemminvert.bc,lemminvert$manag)
anova(lemminvert.disp)
# anosim on original abundances
lemminvert.ano <- anosim(lemminvert.bc, lemminvert$manag,permutations=999)
summary(lemminvert.ano)
# permanova on original abundances
lemminvert.ado <- adonis2(lemminvert.bc~manag,data=lemminvert,permutations=999)
print(lemminvert.ado)
# mv-abund on original abundances
lemminvertmv <- mvabund(lemminvert1)
lemminvertmv.mv <- manyglm(lemminvertmv~lemmenv$manag,family="negative_binomial")
plot(lemminvertmv.mv)
anova(lemminvertmv.mv)
```

Repeat process for proportional abundance data

```{r error=TRUE}
# check homog of dispersions on prop abundances
lemminvert1s.disp <- betadisper(lemminvert1s.bc,lemminvert$manag)
anova(lemminvert1s.disp)
# anosim on proportional abundances
lemminvert1s.ano <- anosim(lemminvert1s.bc, lemminvert$manag,permutations=999)
summary(lemminvert1s.ano)
# permanova on proportional abundances
lemminvert1s.ado <- adonis2(lemminvert1s.bc~manag,data=lemminvert,permutations=999)
print(lemminvert1s.ado)
```

mv-abund on proportional abundances using composition=TRUE (this either doesnt work or is very slow!!!) Note: when composition=TRUE, gives error message; runs with composition=FALSE, but that isn't relative abundance

```{r error=TRUE}
lemminvert1smv.mv <- manyglm(lemminvertmv~lemmenv$manag,composition=TRUE,family="negative_binomial")
plot(lemminvert1smv.mv)
anova(lemminvert1smv.mv)
```

try mv-abund on standardized abunds from earlier

```{r}
lemminvert1smv <- mvabund(lemminvert1s)
lemminvert1smv.mv <- manyglm(lemminvert1smv~lemmenv$manag,family="negative_binomial")
plot(lemminvert1smv.mv)
anova(lemminvert1smv.mv)
```

Finally, run analysis on presence-absence data

```{r}
# check homog of dispersions on pa
lemminvert.jac.disp <- betadisper(lemminvert.jac,lemminvert$manag)
anova(lemminvert.jac.disp)
# anosim on pa
lemminvert.jac.ano <- anosim(lemminvert.jac, lemminvert$manag,permutations=999)
summary(lemminvert.jac.ano)
# permanova on pa
lemminvert.jac.ado <- adonis2(lemminvert.jac~manag,data=lemminvert,permutations=999)
print(lemminvert.jac.ado)
```

mv-abund on pa

```{r}
lemminvert1pa <- decostand(lemminvert1, method = "pa")
lemminvert1pamv <- mvabund(lemminvert1pa)
lemminvert1pamv.mv <- manyglm(lemminvert1pamv~lemmenv$manag,family="binomial")
plot(lemminvert1pamv.mv)
anova(lemminvert1pamv.mv)
```

### Use bio-env to explore relationships with environmental variables

```{r}
# bio-env raw data
bioenv(lemminvert.bc, lemmenv1, method="spearman", metric="euclidean")
# bio-env prop data
bioenv(lemminvert1s.bc, lemmenv1,method="spearman", metric="euclidean")
# bio-env on pa data
bioenv(lemminvert.jac, lemmenv1,method="spearman", metric="euclidean")
# mantel test
lemmenv.euc <- vegdist(lemmenv1,"euclidean")
mantel(lemminvert.bc,lemmenv.euc,method="spearman")
mantel(lemminvert.bc,lemmenv.euc,method="pearson")
```

### model/test continuous variables with permanova and mv-abund

```{r}
# raw abundances
lemminvert2.ado <- adonis2(lemminvert.bc~depth+silt+ph+temp+surface+cond+lchla+tn+tp+reed+emerg,data=lemmenv1,permutations=999)
print(lemminvert2.ado)
lemminvert3.ado <- adonis2(lemminvert.bc~depth+silt+ph+temp+surface+cond+lchla+tn+tp+reed+emerg,data=lemmenv1,by="margin",permutations=999)
print(lemminvert3.ado)
lemminvertmv <- mvabund(lemminvert1)
lemminvertmv.mv <- manyglm(lemminvertmv~lemmenv1$depth+lemmenv1$silt+lemmenv1$ph+lemmenv1$temp+lemmenv1$surface+lemmenv1$cond+lemmenv1$lchla+lemmenv1$tn+lemmenv1$tp+lemmenv1$reed+lemmenv1$emerg,family="negative_binomial")
# use summary to get Wald tests (at least vaguely comparable to Type III Permanova results)
summary(lemminvertmv.mv)
plot(lemminvertmv.mv)
```

### Generate graphs

```{r}
br=c("nm","li","nf","yf")
la=c("None", "Light", "No fish", "Young fish")
a<-as.data.frame(lemminvert.mds$points)
a<-cbind(lemminvert[c(1:3)],a)   #Add site names & symbols from original data file
p1<-ggplot(data=a, aes(x=MDS1, y=MDS2, shape=manag, ) )+
  geom_point()+
  labs(y="MDS2", x="MDS1", title="Raw")+
  scale_shape_manual(values=sym4,
                     name="Management",
                     breaks=br,
                     labels=la,
                     guide =
                         guide_legend(label.theme = element_text(size=6),
                                    title=NULL)
                     )+
  theme_qk()
as<-as.data.frame(lemminvert1s.mds$points)
as<-cbind(lemminvert[c(1:3)],as)   #Add site names & symbols from original data file
p2<-ggplot(data=as, aes(x=MDS1, y=MDS2, shape=manag, ) )+
  geom_point()+
  labs(y=NULL, x="MDS1", title="Standardized")+
  scale_shape_manual(values=sym4,
                     breaks=br,
                     labels=la,
                     guide =
                         guide_legend(label.theme = element_text(size=6),
                                    title=NULL)
                     )+
  theme_qk()
aj<-as.data.frame(lemminvert.jac.mds$points)
aj<-cbind(lemminvert[c(1:3)],aj)   #Add site names & symbols from original data file
p3<-ggplot(data=aj, aes(x=MDS1, y=MDS2, shape=manag, ) )+
  geom_point()+
  labs(y=NULL, x="MDS1", title="Jaccard")+
  scale_shape_manual(values=sym4,
                      breaks=br,
                     labels=la,
                     guide =
                         guide_legend(label.theme = element_text(size=6),
                                    title=NULL)
                     )+
  theme_qk()
p4<-p1+p2+p3+plot_layout(guides='collect')
p4
```

Colour version

```{r}
p1a<-ggplot(data=a, aes(x=MDS1, y=MDS2, color=manag, ) )+
  geom_point()+
  labs(y="MDS2", x="MDS1", title="Raw")+
  scale_color_uchicago(
                      breaks=br,
                     labels=la,
                     guide =
                         guide_legend(label.theme = element_text(size=6),
                                    title=NULL)
                     )+
  theme_qk()
p2a<-ggplot(data=as, aes(x=MDS1, y=MDS2, color=manag, ) )+
  geom_point()+
  labs(y=NULL, x="MDS1", title="Standardized")+
  scale_color_uchicago(
                     name="Management",
                     breaks=br,
                     labels=la,
                     guide =
                         guide_legend(label.theme = element_text(size=6),
                                    title=NULL)
                     )+
  theme_qk()
p3a<-ggplot(data=aj, aes(x=MDS1, y=MDS2, color=manag, ) )+
  geom_point()+
  labs(y=NULL, x="MDS1", title="Jacard")+
  scale_color_uchicago(
                      breaks=br,
                     labels=la,
                     guide =
                         guide_legend(label.theme = element_text(size=6),
                                    title=NULL)
                     )+
  theme_qk()
p4a<-p1a+p2a+p3a+plot_layout(guides='collect')&theme(axis.text=element_blank())
p4a

```
