---
title: "QK Box 2.5"
output:
  html_document:
    df_print: paged
  html_notebook:
    theme: flatly
editor_options:
  chunk_output_type: console
execute: 
  cache: true
knitr:
  opts_chunk: 
    collapse: true
    echo: true
    warning: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)

# Set working directory
knitr::opts_knit$set(root.dir = "../")
```

This box continues with the Low et al. example starting in Box 2.2

### Preliminaries

Use rstanarm and BayesFactor packages; also needs bayestestR.

Add bayesplot for control over plot

```{r}
#| echo: false
#| include: true
#| results: hide

library(tidyverse)
library(rstanarm)
library(BayesFactor)
library(bayestestR)
library(bayesplot)
```

Load graphics packages (if ggplot version of figures wanted)

```{r}
#| echo: false
#| results: hide

source("../R/appearance.R")
```

Note that iso is reference group so diff between means is -ve

### Uninformative priors

```{r}
#| include: true
#| warning: false
#| results: hide
low <- read_csv("../data/lowco2.csv")
low1 <- stan_glm(co2~anesth,family = gaussian(link = "identity"),data=low)
posteriors1 <- describe_posterior(low1)
```

```{r}
print_md(posteriors1, digits = 2)
# plot posterior distribution for all three parameters (intercept, mean diff, sigma)
plot(low1,plotfun="mcmc_hist")
# get Bayes factor for mean diff
lowx <- as.data.frame(low)
lmBF(co2~anesth, data=lowx,posterior=FALSE)
```

### Informative priors

Run three options, mean difference with high and low precision, and a bigger mean difference with high precision

```{r}
#for mean difference with high precision
low2 <- stan_glm(co2~anesth,family = gaussian(link = "identity"),prior=normal(-25,5),data=low)
posteriors2 <- describe_posterior(low2)
print_md(posteriors2, digits = 2)
# informative prior for mean difference with low precision
low3 <- stan_glm(co2~anesth,family = gaussian(link = "identity"),prior=normal(-25,20),data=low)
posteriors3 <- describe_posterior(low3)
print_md(posteriors3, digits = 2)
# informative prior for bigger mean difference with high precision
low4 <- stan_glm(co2~anesth,family = gaussian(link = "identity"),prior=normal(-50,5),data=low)
posteriors4 <- describe_posterior(low4)
print_md(posteriors4, digits = 2)
```

### Generate ggplot-compatible figure for mean difference posterior distribution

```{r}
posterior<-as.array(low1)
color_scheme_set("gray")
p<-mcmc_hist(posterior, pars = c("anesthiso"))+
  xlab("Mean difference")
p
# ggsave ("QK F2_07.pdf", plot = p, height = ph, width = pw, units='cm')

```
