---
title: "QK Box 8.2"

output: 
  html_notebook:
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loyn (1987) selected 56 forest patches in southeastern Victoria, Australia, and related the abundance of forest birds in each patch to six predictor variables: patch area (ha), distance to nearest patch (km), distance to nearest larger patch (km), grazing stock (1 to 5 indicating light to heavy), altitude (m) and years since isolation (years). The aim was to develop a predictive model relating bird abundance to these predictors and to assess which predictors were most important for bird abundance.

![Tawny Frogmouth, Podargus strigoides. Mick Keough, [CC BY 4.0](https://creativecommons.org/licenses/by/4.0/)](../media/frogmouth.jpg)

![Eastern Yellow Robin, *Eopsaltria australis*. Emmet Keough, [CC BY 4.0](https://creativecommons.org/licenses/by/4.0/)](../media/robin.jpg)

This data set was used in the first edition and is available [here](data/loyn.csv)

### Preliminaries

First, load the required packages (car, lm.beta, hier.part)

```{r include=FALSE, results='hide'}
source("../R/libraries.R")   #This is the common library
library(lm.beta)
library(hier.part)
```

Import loyn data file (loyn.csv)

```{r}
loyn <- read.csv("../data/loyn.csv")
head(loyn,10)

```

### Check distributions of predictors and look for correlations

We will use the scatterplotMatrix function from the car package

```{r }
scatterplotMatrix(~abund+area+dist+graze+alt+yearisol, data=loyn, cex=0.33, regLine=FALSE, diagonal=list(method='boxplot'))
```

abund looks symmetrical but two unusual observations for area resulting in non-linear relationships with abund and one outlier for distance to nearest patch

#### Try log(10) transformation of area and dist

```{r }
scatterplotMatrix(~abund+log10(area)+log10(dist)+graze+alt+yearisol, data=loyn, cex=0.33, regLine=FALSE, diagonal=list(method='boxplot'))
```

Plots look much better. Now transform the variables in loyn

```{r }
loyn$logarea <- log10(loyn$area)
loyn$logdist <- log10(loyn$dist)
```

### Now check for collinearity among predictors (using larea and ldist)

Use correlations and VIF

```{r }
scatterplotMatrix(~logarea+logdist+graze+alt+yearisol, data=loyn, cex=0.33, regLine=FALSE, diagonal=list(method='boxplot'))
cor(loyn[,c('logarea','logdist','graze','alt','yearisol')])
vif(lm(abund~ logarea+logdist+graze+alt+yearisol, data=loyn))
```

Collinearity OK so now fit linear model relating response to predictor

```{r }
loyn.lm <- lm(abund~ logarea+logdist+graze+alt+yearisol, data=loyn)
```

### Examine regression diagnostics (residual plot and Cooks D)

```{r }
plot(loyn.lm)
augment(loyn.lm)
```

### Display results of model fitting

Get parameter estimates, their confidence intervals, and tests
```{r }
tidy(loyn.lm, conf.int=TRUE)
```

### Get standarised regression coefficients

Use lm.beta function from lm.beta package

```{r }
lm.beta.loyn <- lm.beta(loyn.lm)
summary(loyn.lm)
tidy(lm.beta.loyn, conf.int=TRUE)
```

Show the partial regression (added-variable) plots

```{r }
avPlots(loyn.lm, ask=F)
```

### Run hierarchical partitioning (used later)

We  need two dataframes that are subsets of loyn containing the response variable and the predictors

```{r}
loyn_abund<-loyn$abund
loyn_pred<-subset(loyn, select = c("logarea","logdist","alt","yearisol","graze"))
```

```{r}
hier.part(loyn_abund, loyn_pred, family="gaussian", gof="Rsqu")
```
