---
title: "Q & K Box 13.5"
output:
    html_notebook:
      theme:flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Morehouse et al. (2016) used genetic analysis to determine the parentage of bear cubs, and cross-classified cubs and their parents as causing problems around humans (see Table 4.2)

The data are [here](../data/morehouse.csv)

Morehouse, A. T., Graves, T. A., Mikle, N. & Boyce, M. S. (2016). Nature vs. Nurture: Evidence for social learning of conflict behaviour in Grizzly Bears. *PLoS One*, 11.

### Preliminaries

First, load the required packages (vcd)

```{r include=FALSE, results='hide'}
library(vcd)
library(lmtest)
library(epitools)
```

Import morehouse data file

```{r}
morehouse <- read.csv("../data/morehouse.csv")
head(morehouse)
```

Tabulate the data

```{r }
morehouse.tab<-xtabs(count~mother + offspring, data=morehouse)
morehouse.tab
```

### Calculate chi-sq. test and odds ratios

Output odds and odds ratio on raw, rather than log, scale

```{r}
chisq.test(morehouse.tab, correct=F)$exp
morehouse.chi<- chisq.test(morehouse.tab, correct=F)
morehouse.chi
morehouse.chi$residuals

#Get odds
lodds(morehouse.tab, log = FALSE)
morehouse.odds <- loddsratio(morehouse.tab, log = FALSE)
summary(morehouse.odds)
confint(morehouse.odds)
```
### Run exact tests

```{r}
fisher.test(morehouse.tab)
oddsratio.midp(morehouse.tab)
```

### Do loglinear model

```{r }
morehouse.glm <- glm(count~mother*offspring, family=poisson, data=morehouse)
summary(morehouse.glm)
anova(morehouse.glm)
morehouse1.glm <- glm(count~mother+offspring, family=poisson, data=morehouse)
summary(morehouse1.glm)
anova(morehouse.glm, morehouse1.glm, test="Chisq")
lrtest(morehouse.glm, morehouse1.glm)
plot(morehouse1.glm)
```
