---
title: "QK Box 16.3"
output:
  html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Parkinson et al. (2020) set tested experimentally the effect of artificial light at night (ALAN) and fish presence on abundances of terrestrial and aquatic invertebrates in the littoral zone of a small lake. They established 20 mesocosms (1 m2 plastic mesh walls and bottom submerged with part of the walls above the water surface) with ten provided with ALAN (solar lights in each corner) and ten controls. Half the mesocosms in each group had fish added, but there were no effects of fish on any variables, so Parkinson et al. ignored fish in the subsequent analyses. After about six weeks, a single pan trap was placed in each mesocosm overnight, and invertebrates were identified to family and counted. We will use these data to do an ordination of mesocosms based on invertebrate family abundances and then fit models to examine differences between the two treatment groups (ALAN vs. no-ALAN).

The paper is [here](https://doi.org/10.1371/journal.pone.0240138)

Parkinson, E., Lawson, J. & Tiegs, S. D. (2020). Artificial light at night at the terrestrial-aquatic interface: Effects on predators and fluxes of insect prey. *PLoS One*, 15, e0240138.

### Preliminaries

```{r results='hide'}
# Load standard set of graphics packages and other libraries
source("../R/appearance.R")
source("../R/libraries.R")
```

Load packages specific to this example: vegan,mvabund

```{r results='hide'}
library(vegan)
library(mvabund)
```

Read in data file ([park2020.csv](../data/park2020.csv)) and remove labels

```{r}
park <- read.csv("../data/park2020.csv")
park1 <- park[,-(1:3)]
```

### MDS on invert abundances

```{r}
park1.bc <- vegdist(park1,'bray')
park1.mds <- metaMDS(park1.bc,k=2,autotransform=FALSE)
stressplot(park1.mds, main="Shepard plot")
park1.mds
ordiplot(park1.mds$points,display="sites",type="n")
points(park1.mds$points,pch=as.numeric(park$symb))
```

With standardisation (wisconsin)

```{r}
park1s <- wisconsin(park1)
park1s.bc <- vegdist(park1s,'bray')   
park1s.mds <- metaMDS(park1s.bc,k=3,autotransform=FALSE,try=40,trymax=80,maxit=200)
stressplot(park1s.mds, main="Shepard plot")
park1s.mds
ordiplot(park1s.mds$points,type="n",display="sites",choices=c(1,3))
points(park1s.mds$points,pch=as.numeric(park$symb))
```

**note** that ordihull might be useful to draw boundaries around each group??

### Check homogeneity of dispersions

Check first with file of abundances, then proportional abundances

```{r}
park1.disp <- betadisper(park1.bc,park$treat)
anova(park1.disp)
park1s.disp <- betadisper(park1s.bc,park$treat)
anova(park1s.disp)
```

### Do anosim on abundances and prop. abundances

```{r}
park1.ano <- anosim(park1.bc, park$treat,permutations=999)
summary(park1.ano)
park1s.ano <- anosim(park1s.bc, park$treat,permutations=999)
summary(park1s.ano)
# do permanova on abundances
park1.ado <- adonis2(park1.bc~treat,data=park,permutations=999)
print(park1.ado)
# do permanova on prop abundances
park1s.ado <- adonis2(park1s.bc~treat,data=park,permutations=999)
print(park1s.ado)
```

### mvabund

Run using negative binomial for raw abundances and double standardized abundances from earlier

```{r}
park1mv <- mvabund(park1)
park1mv.mv <- manyglm(park1mv~park$treat,family="negative.binomial")
plot(park1mv.mv)
anova(park1mv.mv)
park1smv <- mvabund(park1s)
park1smv.mv <- manyglm(park1smv~park$treat,family="negative.binomial")
plot(park1smv.mv)
anova(park1smv.mv)
```

### SIMPER analysis for group differences

Again, abundances then proportional abundances

```{r}
park1.sim <- simper(park1, park$treat, permutations=1000)
summary(park1.sim)
park1s.sim <- simper(park1s, park$treat, permutations=10000)
summary(park1s.sim)
```

### Generate graphs

**Note:** graphs may not run at the moment because of a problem in vegan 2.6.2 relating to the scores function when dealing with mds objects. Expecting a fix soon (as of 7 June 2022). Problem did not exist prior to this version of vegan

```{r}
br=c("ALAN","no ALAN")
la=c("ALAN", "No ALAN")
a<-as.data.frame(scores(park1.mds))
a<-cbind(park[c(1:2)],a)   #Add mesocosms & treatments from original data file
p1<-ggplot(data=a, aes(x=NMDS1, y=NMDS2, shape=treat) )+
  geom_point()+
  labs(y="MDS2", x="MDS1", title="Raw")+
  scale_shape_manual(values=sym2,
                     name="Treat",
                     breaks=br,
                     labels=la,
                     guide =
                         guide_legend(label.theme = element_text(size=6), 
                                      title=NULL)
                     )+
  theme_qk()
as<-as.data.frame(scores(park1s.mds))
as<-cbind(park[c(1:2)],as)   #Add mesocosms & treatments from original data file

p2<-ggplot(data=as, aes(x=NMDS1, y=NMDS2, shape=treat) )+
  geom_point()+
  labs(y=NULL, x="MDS1", title="Standardized")+
  scale_shape_manual(values=sym2,
                     name="Treat",
                     breaks=br,
                     labels=la,
                     guide =
                         guide_legend(label.theme = element_text(size=6),
                                    title=NULL)
                     )+
  theme_qk()

p4<-p1+p2+plot_layout(guides='collect')&theme(axis.text=element_blank())
p4
```

```{r}
# with ellipses
p1a<-p1+stat_ellipse(geom="polygon", alpha= 0.2)
p2a<-p2+stat_ellipse(geom="polygon", alpha= 0.2)
p4a<-p1a+p2a+plot_layout(guides='collect')
p4a

```

Color versions

```{r}
p1<-ggplot(data=a, aes(x=NMDS1, y=NMDS2, color=treat, fill=treat) )+
  geom_point()+
  labs(y="MDS2", x="MDS1", title="Raw")+
  scale_color_uchicago(name="Treat",
                     breaks=br,
                     labels=la,
                     guide =
                         guide_legend(label.theme = element_text(size=6), 
                                      title=NULL)
                     )+
  scale_fill_uchicago(name="Treat",
                     breaks=br,
                     labels=la,
                     guide =
                         guide_legend(label.theme = element_text(size=6),
                                    title=NULL))+
  theme_qk()

p2<-ggplot(data=as, aes(x=NMDS1, y=NMDS2, color=treat, fill=treat) )+
  geom_point()+
  labs(y=NULL, x="MDS1", title="Standardized")+
  scale_color_uchicago(
                     name="Treat",
                     breaks=br,
                     labels=la,
                     guide =
                         guide_legend(label.theme = element_text(size=6),
                                    title=NULL)
                     )+
  scale_fill_uchicago(name="Treat",
                     breaks=br,
                     labels=la,
                     guide =
                         guide_legend(label.theme = element_text(size=6),
                                    title=NULL))+
  theme_qk()

p4<-p1+p2+plot_layout(guides='collect')&theme(axis.text=element_blank())
p4

# with ellipses
p1a<-p1+stat_ellipse(geom="polygon", alpha= 0.2)
p2a<-p2+stat_ellipse(geom="polygon", alpha= 0.2)
p4a<-p1a+p2a+plot_layout(guides='collect')&theme(axis.text=element_blank())
p4a

```
